Lesson 5
========================================================
```{r}
pf = read.csv('./lesson3/pseudo_facebook.tsv', sep = '\t')
```

### Multivariate Data
Notes:

***

### Moira Perceived Audience Size Colored by Age
Notes:

***

### Third Qualitative Variable
Notes:

```{r Third Qualitative Variable}
ggplot(aes(x = gender, y = age),
       data = subset(pf, !is.na(gender))) + geom_boxplot()
```

***

### Plotting Conditional Summaries
Notes:

```{r Plotting Conditional Summaries}
pf.fc_by_age_gender = pf %>%
  filter(!is.na(gender)) %>%
  group_by(age, gender) %>%
  summarise(mean_friend_count = mean(friend_count),
            median_friend_count = median(friend_count),
            n = n())

ggplot(aes(x = age, y = median_friend_count), data = pf.fc_by_age_gender) +
  geom_line(aes(color = gender))
```

***

### Thinking in Ratios
Notes:

***

### Wide and Long Format
Notes:

***

### Reshaping Data
Notes:

```{r}
pf.fc_by_age_gender.wide = dcast(pf.fc_by_age_gender,
                                 age ~ gender,
                                 value.var = 'median_friend_count')
```


***

### Ratio Plot
Notes:

```{r Ratio Plot}
ggplot(aes(x = age, y = female / male), data = pf.fc_by_age_gender.wide) + 
  geom_line() + 
  geom_hline(yintercept = 1, alpha = 0.3, linetype = 2)
```

***

### Third Quantitative Variable
Notes:

```{r Third Quantitative Variable}

```

***

### Cut a Variable
Notes:

```{r Cut a Variable}
pf$year_joined.bucket = cut(pf$year_joined,
                            c(2004,2009,2011,2012,2014))
```

***

### Plotting it All Together
Notes:

```{r Plotting it All Together}
ggplot(aes(x = age, y = friend_count), 
       data = subset(pf, !is.na(year_joined.bucket))) + 
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = median)
```

***

### Plot the Grand Mean
Notes:

```{r Plot the Grand Mean}
ggplot(aes(x = age, y = friend_count), 
       data = subset(pf, !is.na(year_joined.bucket))) + 
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = median) + 
  geom_line(linetype = 2, stat = 'summary', fun.y = mean)
```

***

### Friending Rate
Notes:

```{r Friending Rate}
with(subset(pf,tenure>=1), summary(friend_count/tenure))
```

***

### Friendships Initiated
Notes:

What is the median friend rate?

What is the maximum friend rate?

```{r Friendships Initiated}
ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) + 
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = mean)
```

***

### Bias-Variance Tradeoff Revisited
Notes:

```{r Bias-Variance Tradeoff Revisited}

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary',
            fun.y = mean)

ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

# use geom_smooth instead
ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_smooth(aes(color = year_joined.bucket))

```

***

### Sean's NFL Fan Sentiment Study
Notes:

***

### Introducing the Yogurt Data Set
Notes:

***

### Histograms Revisited
Notes:

```{r Histograms Revisited}
yo = read.csv('./lesson5/yogurt.csv')
yo$id = factor(yo$id)

ggplot(aes(x = price), data = yo) + 
  geom_histogram(binwidth = 1)
```

***

### Number of Purchases
Notes:

```{r Number of Purchases}
yo$all.purchases = yo$strawberry + yo$blueberry + yo$pina.colada + yo$plain + yo$mixed.berry

ggplot(aes(x = all.purchases), data = yo) + 
  geom_histogram()
```

***

### Prices over Time
Notes:

```{r Prices over Time}
ggplot(aes(x = time, y = price), data = yo) + 
  geom_jitter(alpha = 1/4, shape = 21)
```

***

### Sampling Observations
Notes:

***

### Looking at Samples of Households

```{r Looking at Sample of Households}
set.seed(4321)
sample.ids = sample(levels(yo$id), 16)

ggplot(aes(x = time, y = price), data = subset(yo, id %in% sample.ids)) + 
  facet_wrap( ~ id) +
  geom_line() + 
  geom_point(aes(size = all.purchases), pch = 1)
```

***

### The Limits of Cross Sectional Data
Notes:

For each user, there is time series data of each purchase (third dimension of dataset, meaning each user has multiple records along time), therefore, we can use facet to understand the behvior change of each user.

However, in pseudo facebook data, data is a snapshot at one moment, which means only one facet of each user, so facet cannot be used to understand each user.
***

### Many Variables
Notes:

***

### Scatterplot Matrix
Notes:
```{r}
set.seed(1836)
pf_subset = pf[,c(2:15)]  # remove user id
ggpairs(pf_subset[sample.int(nrow(pf_subset),1000),])
```

***

### Even More Variables
Notes:

***

### Heat Maps
Notes:

```{r}
nci <- read.table("nci.tsv")
colnames(nci) <- c(1:64)
```

```{r}
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))
```


***

### Analyzing Three of More Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

